---
title: "milo visualization"
output: html_notebook
---

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(edgeR)
library(igraph)
library(SingleCellExperiment)
library(scran)
library(scater)
library(irlba)
library(ggthemes)
library(ggsci)
library(mvtnorm)
library(umap)
library(dplyr)

# BiocManager::install("cydar")
library(cydar)
library(dyntoy)

# devtools::install_github("analyxcompany/ForceAtlas2")
library(ForceAtlas2)
```

Goal: make a nice visualization of the neighborhood (NH) graph

1. The graph layout must be **easily comparable with the graph layout of the whole single-cell dataset**: 
  - we could do this by using ForceAtlas2 to make a force-directed graph layout of the full KNN graph, and then initialize the layout of the NH-graph from the whole KNN-graph layout (as in `scanpy.pl.paga_compare`). I'm using FA2 implementation for R found (here)[https://github.com/analyxcompany/ForceAtlas2] --> takes 2 million years, thank u next
  - Do the initialization on the UMAP
2. The viz must **convey info about the size of a NH**

## Make dataset

```{r}
simulate_linear_trajectory <- function(n.milestones=3, total.size=100){
    dataset <- generate_dataset(
        model = model_linear(num_milestones = n.milestones),
        num_cells = total.size,
        num_features = 2000
        )
    
    ## Build SingleCellExperiment object
    cnts <- t(dataset$counts)
    coldata <- data.frame(row.names = colnames(cnts), dataset$prior_information$groups_id)
    
    sce <- SingleCellExperiment(assays=list(counts=cnts), colData=coldata)
    mylo <- Milo(sce)
}


sim_data <- simulate_linear_trajectory(n.milestones = 5, total.size = 5000)
sim_data
```
PCA
```{r}
sim_data <- logNormCounts(sim_data)
sim.pca <- prcomp_irlba(t(logcounts(sim_data)), 
                           n=50, center=TRUE, scale.=TRUE)
reducedDim(sim_data, "PCA") <- sim.pca$x
attr(reducedDim(sim_data, "PCA"), "rotation") <- sim.pca$rotation
```

```{r}
sim_data <- buildGraph(sim_data, k=40, d=30, seed=42)
sim_data <- makeNhoods(sim_data, prop=0.25, k=40, d=15, refined=TRUE, reduced_dims="PCA")

# sim_data <- countCells(sim_data, meta.data=thymus.meta, samples="ExpSamp")
```
```{r}
assay(sim_data, "scaled_logcounts") <- t(scale(t(logcounts(sim_data)))) 
sim_data <- calcNhoodExpression(sim_data, assay = "scaled_logcounts")
sim_data <- projectNhoodExpression(sim_data, d = 50, scale = FALSE, center=FALSE)

l2.norm.dims <- t(apply(nhoodReducedDim(sim_data), 1, function(x) x/.l2norm(x)))

## Compute UMAP
nh_index <- str_subset(rownames(nhoodReducedDim(sim_data)), "^nh")
c_index <- setdiff(rownames(nhoodReducedDim(sim_data)), nh_index)


nhood.umap <- uwot::umap(nhoodReducedDim(sim_data)[nh_index, c(1:30)],
                          n_components=2,
                          n_neighbors=30, metric='cosine',
                          init='random', min_dist=0.7)
c.umap <- uwot::umap(l2.norm.dims[c_index, c(1:30)],
                          n_components=2,
                          n_neighbors=30, metric='cosine',
                          init='random', min_dist=0.7, ret_model = TRUE)

nhood.umap <- uwot::umap_transform(l2.norm.dims[nh_index, c(1:30)], model = c.umap)


# nhoodReducedDim(sim_data, "UMAP") <- nhood.umap$layout
data.frame(c.umap$embedding) %>%
  rownames_to_column() %>%
  mutate(class=ifelse(str_detect(rowname, "^nh_"), "nh", "cell")) %>%
  ggplot(aes(X1, X2)) + geom_point(color="grey") +
  geom_point(data=. %>% filter(class=='nh'))

data.frame(nhood.umap) %>%
  mutate(class="nh") %>%
  # bind_rows(data.frame(c.umap$embedding) %>% mutate(class="cell")) %>%
  # rownames_to_column() %>%
  # mutate(class=ifelse(str_detect(rowname, "^nh_"), "nh", "cell")) %>%
  ggplot(aes(X1, X2)) + geom_point(color="grey") +
  geom_point(data=. %>% filter(class=='nh'))


```

```{r}
#' Calculates the l2-norm of a vector
#'
#' Adapted from PMA package
#' @references Witten, Tibshirani, and Hastie, Biostatistics 2009
#' @references \url{https://github.com/cran/PMA/blob/master/R/PMD.R}
#'
#' @param vec numeric vector
#'
#' @return returns the l2-norm.
#'
.l2norm <- function(vec) {
  a <- sqrt(x = sum(vec ^ 2))
  if (a == 0) {
    a <- .05
  }
  return(a)
}

## Project profiles to same PCA space of the single-cells
X_reduced_dims <- reducedDim(sim_data, "PCA")
loadings <- attr(X_reduced_dims, "rotation")
n.reducedDim <- t(scale(nhoodExpression(sim_data), scale=TRUE, center=TRUE)) %*% loadings
n.reducedDim <- n.reducedDim[, c(1:30)]

## Make one PC matrix including single-cells and nhoods
rownames(n.reducedDim) <-
    paste0("nh_", seq(1:nrow(n.reducedDim)))
X_reduced_dims_merged <- rbind(n.reducedDim, X_reduced_dims[, c(1:30)])

## Do l2 norm


data.frame(t(apply(X_reduced_dims_merged, 1, function(x) x/.l2norm(x)))) %>%
  rownames_to_column() %>%
  mutate(class=ifelse(str_detect(rowname, "^nh_"), "nh", "cell")) %>%
  ggplot(aes(PC1, PC2)) + geom_point(color="grey") +
  geom_point(data=. %>% filter(class=='nh')) +
  facet_wrap(class~.)
```

```{r}
hist(nhoodExpression(sim_data)[100,], breaks=100)
hist(assay(sim_data, "scaled_logcounts")[100,], breaks=100)
```
```{r}
merged_cnts <- cbind(nhoodExpression(sim_data), logcounts(sim_data))

## Scale data
merged_cnts_scaled <- t(scale(t(merged_cnts)))

## Run PCA
merged_pca <- prcomp_irlba(t(merged_cnts_scaled), n = 30, center=FALSE, scale. = FALSE)
# pca_mat <- rbind(merged_pca$x[(ncol(liver_milo)+1):(ncol(liver_milo)+(length(nhoods(liver_milo)))),], merged_pca$x[colnames(liver_milo),])
## Add to slot nhoodsReducedDim
nhoodReducedDim(sim_data, "PCA") <- merged_pca$x

## Run UMAP on joint PCA
umap_out <- uwot::umap(nhoodReducedDim(sim_data, "PCA"), n_neighbors = 50, n_components = 2, scale=FALSE)

```

```{r}
data.frame(umap_out, row.names = rownames(nhoodReducedDim(sim_data))) %>%
  # mutate(class="nh") %>%
  # bind_rows(data.frame(c.umap$embedding) %>% mutate(class="cell")) %>%
  rownames_to_column() %>%
  mutate(class=ifelse(as.numeric(rowname) < length(nhoods(sim_data)), "nh", "cell")) %>%
  ggplot(aes(X1, X2, color=class)) + geom_point(color="grey") +
  geom_point(data=. %>% filter(class=='nh'), size=0.1)

```

## Thymus mouse dataset
```{r}
th_milo <- readRDS("~/MouseThymus_milo.RDS")

data.frame(t(apply(nhoodReducedDim(th_milo), 1, function(x) x/.l2norm(x)))) %>%
# data.frame(nhoodReducedDim(th_milo)) %>%  
  rownames_to_column() %>%
  mutate(class=ifelse(str_detect(rowname, "^nh_"), "nh", "cell")) %>%
  ggplot(aes(PC1, PC2)) + geom_point(color="grey") +
  geom_point(data=. %>% filter(class=='nh')) +
  facet_wrap(class~.)

umap_1 <- uwot::umap(nhoodReducedDim(th_milo)[,c(1:15)],
                          n_components=2,
                          n_neighbors=30, metric='cosine',
                          init='random', min_dist=0.7)

umap_2 <- uwot::umap(t(apply(nhoodReducedDim(th_milo), 1, function(x) x/.l2norm(x)))[,c(1:15)],
                          n_components=2,
                          n_neighbors=30, metric='cosine',
                          init='random', min_dist=0.7)

```

```{r}
data.frame(umap_1) %>%
# data.frame(nhoodReducedDim(th_milo)) %>%  
  rownames_to_column() %>%
  mutate(class=ifelse(as.numeric(rowname) < length(nhoods(th_milo)), "nh", "cell")) %>%
  ggplot(aes(X1, X2, color=class)) + geom_point(color="grey") +
  geom_point(data=. %>% filter(class=='nh')) +
  ggtitle("project")

data.frame(umap_2) %>%
# data.frame(nhoodReducedDim(th_milo)) %>%  
  rownames_to_column() %>%
  mutate(class=ifelse(as.numeric(rowname) < length(nhoods(th_milo)), "nh", "cell")) %>%
  ggplot(aes(X1, X2, color=class)) + geom_point(color="grey") +
  geom_point(data=. %>% filter(class=='nh'))

```
```{r}
merged_cnts <- cbind(nhoodExpression(th_milo), logcounts(th_milo)[rownames(nhoodExpression(th_milo)),])

## Scale data
merged_cnts_scaled <- t(scale(t(merged_cnts)))

## Run PCA
merged_pca <- prcomp_irlba(t(merged_cnts_scaled), n = 30, center=FALSE, scale. = FALSE)
# pca_mat <- rbind(merged_pca$x[(ncol(liver_milo)+1):(ncol(liver_milo)+(length(nhoods(liver_milo)))),], merged_pca$x[colnames(liver_milo),])
## Add to slot nhoodsReducedDim
nhoodReducedDim(th_milo, "PCA") <- merged_pca$x

## Run UMAP on joint PCA
umap_3 <- uwot::umap(nhoodReducedDim(th_milo, "PCA"), n_neighbors = 50, n_components = 2, scale=FALSE)

```

```{r}
data.frame(umap_3, row.names = rownames(nhoodReducedDim(th_milo))) %>%
  # mutate(class="nh") %>%
  # bind_rows(data.frame(c.umap$embedding) %>% mutate(class="cell")) %>%
  rownames_to_column() %>%
  mutate(class=ifelse(as.numeric(rowname) < length(nhoods(th_milo)), "nh", "cell")) %>%
  ggplot(aes(X1, X2, color=class)) + geom_point(color="grey") +
  geom_point(data=. %>% filter(class=='nh'), size=0.6)

```

```{r}
umap_1 <- uwot::umap(reducedDim(th_milo, "PCA")[,c(1:15)],
                          n_components=2,
                          n_neighbors=30, metric='cosine',
                          init='random', min_dist=0.7)

reducedDim(th_milo, "UMAP") <- umap_1

plotReducedDim(th_milo, "UMAP", colour_by = "Cluster")
```
```{r}
data.frame(colData(th_milo)[unlist(nhoodIndex(th_milo)),]) %>%
  mutate(nhood=unlist(nhoodIndex(th_milo))) %>%
  filter(Cluster=="nTEC") 

nh_intersect_mat["3","857"]
```

Make graph abstraction
```{r}
l <- nhoods(th_milo)

nms <- gtools::permutations(n=length(names(l)), v = names(l), r = 2,repeats.allowed = T)
ll <- gtools::permutations(n=length(l), v = l, r = 2,repeats.allowed = T)

out2 <- sapply( 1:nrow(nms) , function(x) length( intersect( nhoods(th_milo)[[nms[x,1]]] ,nhoods(th_milo)[[ nms[x,2] ]]) ) )

nh_intersect_mat <-
  data.frame(nh1=nms[,1], nh2=nms[,2], int_len=out2) %>%
  pivot_wider(id_cols = nh1, names_from=nh2, values_from=int_len) %>%
  column_to_rownames('nh1') %>%
  as.matrix()

pheatmap::pheatmap(scale_to_a(nh_intersect_mat_filt))
```

```{r}
plotReducedDim(th_milo, "UMAP", colour_by = "Cluster")
```


```{r, fig.width=15, fig.height=15}
scale_to_a <- function(x){
  (x - min(x))/(max(x) - min(x))
}

## Keep as edges only 5 or more
nh_intersect_mat_filt <- ifelse(nh_intersect_mat < 5, 0, nh_intersect_mat)
ig <- graph.adjacency(nh_intersect_mat, mode="undirected", weighted=TRUE)
# graph <- igraph::graph_from_adjacency_matrix(nh_intersect_mat, mode = "undirected", weighted = TRUE)

ig <- set_vertex_attr(ig, name = 'size', value = sapply(nhoods(th_milo)[as.numeric(vertex.attributes(ig)$name)], length))
min_size = 2

plot.igraph(simplify(ig),
            # vertex.size=3, 
            arrow.size=0, 
            layout=reducedDim(th_milo[,as.numeric(vertex.attributes(ig)$name)], "UMAP"),
            edge.width=scale_to_a(edge.attributes(ig)$weight), 
            vertex.size=scale_to_a(vertex.attributes(ig)$size)*5 + min_size , 
            vertex.label=NA, vertex.border=0)

hist(scale_to_a(edge.attributes(ig)$weight))
```



---

## Old code / junk

```{r}
dataset <- generate_dataset(
    model = model_linear(num_milestones = 5),
    num_cells = 5000,
    num_features = 2000
  )

## Build SingleCellExperiment object
cnts <- t(dataset$counts)
coldata <- data.frame(row.names = colnames(cnts), dataset$prior_information$groups_id)

sce <- SingleCellExperiment(assays=list(counts=cnts), colData=coldata)

## Add KNN graph
pca <- prcomp_irlba(t(cnts), n=30, scale.=TRUE, center=TRUE)
X_pca <- pca$x[, c(1:30)]
knn_graph <- buildKNNGraph(x=X_pca, k=21, d=NA, transposed=TRUE)
X_umap <- umap(X_pca, n_components=2, n_neighbors=21, metric='euclidean',init='random', min_dist=0.1)
colnames(X_umap$layout) <- c("UMAP_1", "UMAP_2")

reducedDims(sce) <- list(PCA=X_pca, UMAP=X_umap$layout)
```

```{r}
plot(reducedDim(sce, 'UMAP')[,1], reducedDim(sce, 'UMAP')[,2])
```

<!-- Build the NHs -->

<!-- ```{r} -->
<!-- neighborhood_expression <- function(neighborhoods, data.set){ -->
<!--   # I'll calculate the average value of each neighborhood for each of the n features in the data.set -->
<!--   neighbour.model <- matrix(0L, ncol=length(neighborhoods), nrow=ncol(data.set)) -->
<!--   # neighbour.model <- sapply(1:length(neighborhoods), FUN=function(X) print(length(neighbour.model[, X]))) -->
<!--   for(x in seq_along(1:length(neighborhoods))){ -->
<!--     neighbour.model[neighborhoods[[x]], x] <- 1 -->
<!--   } -->

<!--   neigh.exprs <- t(neighbour.model) %*% t(data.set) -->
<!--   neigh.exprs <- t(apply(neigh.exprs, 2, FUN=function(XP) XP/colSums(neighbour.model))) -->
<!--   return(neigh.exprs) -->
<!-- } -->

<!-- ## Pick random centers  -->
<!-- random.vertices <- sample(V(knn_graph), size=floor(0.1*length(V(knn_graph)))) -->
<!-- vertex.list <- sapply(1:length(random.vertices), FUN=function(X) neighbors(knn_graph, v=random.vertices[X])) -->

<!-- ## Compute avg expression profile within neighborhood -->
<!-- nh_cnts <- neighborhood_expression(vertex.list, cnts) -->

<!-- ## Make graph -->
<!-- nh_pca <- prcomp_irlba(t(nh_cnts), n=30, scale.=TRUE, center=TRUE) -->
<!-- nh_knn_graph <- buildKNNGraph(x=nh_pca$x, k=21, d=NA, transposed=TRUE) -->
<!-- nh_umap_random <- umap(nh_pca$x, n_components=2, n_neighbors=21, metric='euclidean',init='random', min_dist=0.1) -->
<!-- ``` -->

<!-- Take initial UMAP coordinates from the position in the original UMAP -->

<!-- ```{r} -->
<!-- plot(reducedDim(sce[,as.vector(random.vertices)], "UMAP")[,1], reducedDim(sce[,as.vector(random.vertices)], "UMAP")[,2]) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- nh_umap_init <- umap(nh_pca$x, n_components=2, n_neighbors=21, metric='euclidean',init=reducedDim(sce[,as.vector(random.vertices)], "UMAP"), min_dist=0.1) -->

<!-- plot(nh_umap_init$layout[,1], nh_umap_init$layout[,2]) -->
<!-- ``` -->

<!-- This doesn't work. What if we cheat and we just put spots over the UMAP? -->
<!-- ```{r} -->
<!-- if (!"Vertex" %in% colnames(colData(sce))) { -->
<!--   colData(sce) <- as.data.frame(colData(sce)) %>% -->
<!--     rowid_to_column("Vertex")  -->
<!--   }  -->
<!-- colData(sce) <- colData(sce)[c("Vertex", "cell_id", "group_id")] -->
<!-- nh_size <- sapply(vertex.list, function(x) length(x)) -->
<!-- nh_df <- data.frame(Vertex=as.vector(random.vertices), nh_size) -->

<!-- colData(sce) <- as.data.frame(colData(sce)) %>% -->
<!--   left_join(nh_df, by="Vertex") %>% -->
<!--   dplyr::mutate(is_nh=ifelse(is.na(nh_size),FALSE, TRUE)) %>% -->
<!--   DataFrame() -->

<!-- data.frame(reducedDim(sce, 'UMAP'), colData(sce)) %>% -->
<!--   dplyr::filter(is_nh) %>% -->
<!--   ggplot(aes(UMAP_1, UMAP_2)) + -->
<!--   geom_point(aes(size=nh_size), alpha=0.1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- data.frame(colData(sce)) %>% -->
<!--   dplyr::filter(is_nh) %>% -->
<!--   dplyr::mutate(UMAP_1=nh_umap_init$layout[,1], UMAP_2=nh_umap_init$layout[,2]) %>% -->
<!--   ggplot(aes(UMAP_1, UMAP_2)) + -->
<!--   geom_point( -->
<!--     aes(size=nh_size), -->
<!--     alpha=0.1) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- # ## Calculate overlap between nh -->
<!-- # # Get the combinations of names of list elements -->
<!-- # l <- vertex.list -->
<!-- # nms <- combn(seq_along(vertex.list) , 2 , FUN = paste0 , collapse = "_" , simplify = FALSE ) -->
<!-- #  -->
<!-- # # Make the combinations of list elements -->
<!-- # ll <- combn( l , 2 , simplify = FALSE ) -->
<!-- #  -->
<!-- # # Intersect the list elements -->
<!-- # out <- sapply( ll , function(x) length( intersect( x[[1]] , x[[2]] ) ) ) -->
<!-- # data.frame(pair=unlist(nms), overlap=out) -->

<!-- ``` -->


