---
title: "Milo scalability"
output: html_notebook
---

I have simulated data on upto 200k cells from a single linear trajectory. Using these data I 
have performed the various steps in `Milo`, up to and including differential abundance testing 
and recorded the clock time using different numbers of cells. The results are plotted here.

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggthemes)
library(ggsci)
library(scales)
library(reshape2)
library(ggrepel)
```


```{r}
# timing.dir <- "~/Dropbox/Milo/simulations/timing.dir/"
# time.files <- list.files(timing.dir, full.names = TRUE)
# time.list <- list()
# 
# for(x in seq_along(time.files)){
#     x.time <- read.table(time.files[x], sep="\t", header=TRUE, stringsAsFactors=FALSE)
#     time.list[[paste0(x)]] <- x.time
# }

# time.df <- do.call(rbind.data.frame, time.list)
time.df <- read.table("~/Dropbox/Milo/timing.dir/Trajectory_simulation-timing.tsv",
                      sep="\t", header=TRUE, stringsAsFactors=FALSE)
time.df$DataSet <- "Simulation"
```


```{r, warning=FALSE, message=FALSE}
ggplot(time.df, aes(x=NCells, y=Elapsed/60)) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Elapsed time (mins)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Simulation_timing.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Simulation_timing.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
memory.df <- read.table("~/Dropbox/Milo/timing.dir/Trajectory_simulation-MemProfile.tsv",
                        sep="\t", header=TRUE, stringsAsFactor=FALSE)

memory.df$Total.Mem <- apply(memory.df[, c(1:6)], 1, sum)
memory.df$DataSet <- "Simulation"
```

I've profiled the memory usage for each function in the Milo workflow.

```{r, warning=FALSE, message=FALSE}
ggplot(memory.df, aes(x=NCells, y=Mem.Graph/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Simulation_memory-GraphBuilding.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Simulation_memory-GraphBuilding.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```

```{r, warning=FALSE, message=FALSE}
ggplot(memory.df, aes(x=NCells, y=Mem.Nhoods/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Simulation_memory-makeNhoods.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Simulation_memory-makeNhoods.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(memory.df, aes(x=NCells, y=Mem.Distance/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Simulation_memory-CalcNhoodDistance.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Simulation_memory-CalcNhoodDistance.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(memory.df, aes(x=NCells, y=Mem.Count/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Simulation_memory-countCells.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Simulation_memory-countCells.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(memory.df, aes(x=NCells, y=Mem.Test/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Simulation_memory-testNhoods.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Simulation_memory-testNhoods.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(memory.df, aes(x=NCells, y=Total.Mem/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Simulation_memory-TotalMem.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Simulation_memory-TotalMem.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```

## Liver data scalability

```{r}
liver.time.df <- read.table("~/Dropbox/Milo/timing.dir/Liver-Timing.tsv",
                            sep="\t", header=TRUE, stringsAsFactors=FALSE)
liver.time.df$DataSet <- "Liver"
```


```{r, warning=FALSE, message=FALSE}
ggplot(liver.time.df, aes(x=NCells, y=Elapsed/60)) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Elapsed time (mins)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Liver_timing.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Liver_timing.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
liver.memory.df <- read.table("~/Dropbox/Milo/timing.dir/Liver-MemProfile.tsv",
                              sep="\t", header=TRUE, stringsAsFactor=FALSE)

liver.memory.df$Total.Mem <- apply(liver.memory.df[, c(1:6)], 1, sum)
liver.memory.df$DataSet <- "Liver"
```

I've profiled the memory usage for each function in the Milo workflow.

```{r, warning=FALSE, message=FALSE}
ggplot(liver.memory.df, aes(x=NCells, y=Mem.Graph/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Liver_memory-GraphBuilding.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Liver_memory-GraphBuilding.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```

```{r, warning=FALSE, message=FALSE}
ggplot(liver.memory.df, aes(x=NCells, y=Mem.Nhoods/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Liver_memory-makeNhoods.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Liver_memory-makeNhoods.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(liver.memory.df, aes(x=NCells, y=Mem.Distance/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Liver_memory-CalcNhoodDistance.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Liver_memory-CalcNhoodDistance.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(liver.memory.df, aes(x=NCells, y=Mem.Count/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Liver_memory-countCells.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Liver_memory-countCells.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(liver.memory.df, aes(x=NCells, y=Mem.Test/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Liver_memory-testNhoods.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Liver_memory-testNhoods.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(liver.memory.df, aes(x=NCells, y=Mem.Obj/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Liver_memory-objectSize.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Liver_memory-objectSize.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(liver.memory.df, aes(x=NCells, y=Total.Mem/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Liver_memory-TotalMem.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Liver_memory-TotalMem.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


## Thymus data scalability

```{r}
thymus.time.df <- read.table("~/Dropbox/Milo/timing.dir/Thymus-Timing.tsv",
                            sep="\t", header=TRUE, stringsAsFactors=FALSE)
thymus.time.df$DataSet <- "Thymus"
```


```{r, warning=FALSE, message=FALSE}
ggplot(thymus.time.df, aes(x=NCells, y=Elapsed/60)) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Elapsed time (mins)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Thymus_timing.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Thymus_timing.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
thymus.memory.df <- read.table("~/Dropbox/Milo/timing.dir/Thymus-MemProfile.tsv",
                              sep="\t", header=TRUE, stringsAsFactor=FALSE)

thymus.memory.df$Total.Mem <- apply(thymus.memory.df[, c(1:6)], 1, sum)
thymus.memory.df$DataSet <- "Thymus"
```

I've profiled the memory usage for each function in the Milo workflow.

```{r, warning=FALSE, message=FALSE}
ggplot(thymus.memory.df, aes(x=NCells, y=Mem.Graph/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Thymus_memory-GraphBuilding.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Thymus_memory-GraphBuilding.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```

```{r, warning=FALSE, message=FALSE}
ggplot(thymus.memory.df, aes(x=NCells, y=Mem.Nhoods/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Thymus_memory-makeNhoods.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Thymus_memory-makeNhoods.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(thymus.memory.df, aes(x=NCells, y=Mem.Distance/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Thymus_memory-CalcNhoodDistance.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Thymus_memory-CalcNhoodDistance.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(thymus.memory.df, aes(x=NCells, y=Mem.Count/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Thymus_memory-countCells.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Thymus_memory-countCells.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(thymus.memory.df, aes(x=NCells, y=Mem.Test/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Thymus_memory-testNhoods.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Thymus_memory-testNhoods.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(thymus.memory.df, aes(x=NCells, y=Mem.Obj/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Thymus_memory-objectSize.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Thymus_memory-objectSize.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(thymus.memory.df, aes(x=NCells, y=Total.Mem/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Thymus_memory-TotalMem.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Thymus_memory-TotalMem.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```

## Gastrulation data scalability

```{r}
gast.time.df <- read.table("~/Dropbox/Milo/timing.dir/Gastrulation-Timing.tsv",
                            sep="\t", header=TRUE, stringsAsFactors=FALSE)
gast.time.df$DataSet <- "Gastrulation"
```


```{r, warning=FALSE, message=FALSE}
ggplot(gast.time.df, aes(x=NCells, y=Elapsed/60)) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Elapsed time (mins)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Gastrulation_timing.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Gastrulation_timing.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
gast.memory.df <- read.table("~/Dropbox/Milo/timing.dir/Gastrulation-MemProfile.tsv",
                              sep="\t", header=TRUE, stringsAsFactor=FALSE)

gast.memory.df$Total.Mem <- apply(gast.memory.df[, c(1:6)], 1, sum)
gast.memory.df$DataSet <- "Gastrulation"
```

I've profiled the memory usage for each function in the Milo workflow.

```{r, warning=FALSE, message=FALSE}
ggplot(gast.memory.df, aes(x=NCells, y=Mem.Graph/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-GraphBuilding.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-GraphBuilding.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```

```{r, warning=FALSE, message=FALSE}
ggplot(gast.memory.df, aes(x=NCells, y=Mem.Nhoods/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-makeNhoods.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-makeNhoods.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(gast.memory.df, aes(x=NCells, y=Mem.Distance/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-CalcNhoodDistance.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-CalcNhoodDistance.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(gast.memory.df, aes(x=NCells, y=Mem.Count/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-countCells.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-countCells.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(gast.memory.df, aes(x=NCells, y=Mem.Test/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-testNhoods.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-testNhoods.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(gast.memory.df, aes(x=NCells, y=Mem.Obj/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-objectSize.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-objectSize.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(gast.memory.df, aes(x=NCells, y=Total.Mem/(1024))) +
    geom_point(size=3) +
    theme_bw() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'))

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-TotalMem.png",
       height=4.15, width=4.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/Gastrulation_memory-TotalMem.pdf",
       height=4.15, width=4.75, useDingbats=FALSE)
```



## Combining across data sets

```{r, warning=FALSE , message=FALSE}
all.timing <- do.call(rbind.data.frame, 
                      list("Sim"=time.df[time.df$NCells %in% max(time.df$NCells), , drop=FALSE], 
                           "liver"=liver.time.df[liver.time.df$NCells %in% max(liver.time.df$NCells), , drop=FALSE], 
                           "thymus"=thymus.time.df[thymus.time.df$NCells %in% max(thymus.time.df$NCells), , drop=FALSE],
                           "gast"=gast.time.df[gast.time.df$NCells %in% max(gast.time.df$NCells), , drop=FALSE]))

common.cols <- intersect(colnames(memory.df), colnames(liver.memory.df))
all.memory <- do.call(rbind.data.frame, list("Sim"=memory.df[memory.df$NCells %in% max(memory.df$NCells), common.cols], 
                                             "liver"=liver.memory.df[liver.memory.df$NCells %in% max(liver.memory.df$NCells), common.cols],
                                             "thymus"=thymus.memory.df[thymus.memory.df$NCells %in% max(thymus.memory.df$NCells), common.cols],
                                             "gast"=gast.memory.df[gast.memory.df$NCells %in% max(gast.memory.df$NCells), common.cols]))
```


```{r, warning=FALSE , message=FALSE}
down.timing <- do.call(rbind.data.frame, 
                      list("Sim"=time.df[!time.df$NCells %in% max(time.df$NCells), , drop=FALSE], 
                           "liver"=liver.time.df[!liver.time.df$NCells %in% max(liver.time.df$NCells), , drop=FALSE], 
                           "thymus"=thymus.time.df[!thymus.time.df$NCells %in% max(thymus.time.df$NCells), , drop=FALSE],
                           "gast"=gast.time.df[!gast.time.df$NCells %in% max(gast.time.df$NCells), , drop=FALSE]))


down.memory <- do.call(rbind.data.frame, list("Sim"=memory.df[!memory.df$NCells %in% max(memory.df$NCells), common.cols], 
                                              "liver"=liver.memory.df[!liver.memory.df$NCells %in% max(liver.memory.df$NCells), common.cols],
                                              "thymus"=thymus.memory.df[!thymus.memory.df$NCells %in% max(thymus.memory.df$NCells), common.cols],
                                              "gast"=gast.memory.df[!gast.memory.df$NCells %in% max(gast.memory.df$NCells), common.cols]))
```




```{r, warning=FALSE, message=FALSE}
ggplot(all.timing, aes(x=NCells, y=Elapsed/60, 
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.timing, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Elapsed time (mins)") +
    # scale_x_log10() +
    # scale_y_log10() +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round(Elapsed/60, 2)), colour='black', size=5,
                    point.padding=0.5,
                    force=25)

ggsave("~/Dropbox/Milo/figures/All_timing.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_timing.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```




```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=NCells, y=Total.Mem/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round(Total.Mem/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=NCells, y=(Total.Mem-Mem.Obj)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round((Total.Mem-Mem.Obj)/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=N.Nhoods, y=(Total.Mem-Mem.Obj)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Neighbourhoods", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round((Total.Mem-Mem.Obj)/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-byNhood.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-byNhood.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```

Is this about the neighbourhood size? I didn't record the average nhood size for my sins.


```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=Median.Nhood.Size, y=(Total.Mem-Mem.Obj)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="Median Nhood Size", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round((Total.Mem-Mem.Obj)/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-byNhoodSize.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-byNhoodSize.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=NCells, y=(Mem.Graph)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round(Mem.Graph/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-buildGraph.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-buildGraph.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```



```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=NCells, y=(Mem.Nhoods)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round(Mem.Nhoods/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-makeNhoods.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-makeNhoods.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=NCells, y=(Mem.Distance)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round(Mem.Distance/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-calcNhoodDist.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-calcNhoodDist.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=NCells, y=(Mem.Count)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round(Mem.Count/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-countCells.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-countCells.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=NCells, y=(Mem.Test)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round(Mem.Test/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-testNhoods.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem_sansObj-testNhoods.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```


```{r, warning=FALSE, message=FALSE}
ggplot(all.memory, aes(x=NCells, y=(Mem.Obj)/(1024),
                       shape=DataSet, colour=DataSet)) +
    geom_point(data=down.memory, size=3,
               colour='grey80', alpha=0.5) +
    geom_point(size=5) +
    theme_bw() +
    scale_colour_npg() +
    labs(x="#Cells", y="Total memory (MB)") +
    theme(axis.text=element_text(size=16, colour='black'),
          axis.title=element_text(size=18, colour='black'),
          legend.title=element_text(size=18, colour='black'),
          legend.text=element_text(size=16, colour='black')) +
    guides(colour=guide_legend(override.aes=list(size=5))) +
    geom_text_repel(aes(label=round(Mem.Obj/(1024), 1)), colour='black', size=5,
                    point.padding=0.5,
                    force=55)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem-ObjSize.png",
       height=4.15, width=6.75, dpi=300)

ggsave("~/Dropbox/Milo/figures/All_memory-TotalMem-ObjSize.pdf",
       height=4.15, width=6.75, useDingbats=FALSE)
```



