---
title: "Mouse thymus: Validating differential abundance findings"
output: html_notebook
---

# Introduction

I have integrated the SMART-seq and droplet mouse thymus data sets. Now we can examine whether the proportions of the cells in the droplet data 
vary with age, and if the most similar cells to the depleted mTEC-biased Intertypical TEC are also depleted with age.

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(reshape2)
library(ggthemes)
library(cowplot)
library(ggrastr)
library(ggsci)
library(RColorBrewer)
library(scales)
library(viridis)
```

What does the integrated UMAP look like?

```{r, warning=FALSE, message=FALSE}
umap.df <- read.table("~/Dropbox/Milo/simulations/Combined_UMAP_meta.tsv",
                      sep="\t", header=TRUE, stringsAsFactors=FALSE)
```

```{r, warning=FALSE, message=FALSE, fig.height=2.55, fig.width=6.15}
umap.df$Dataset <- factor(umap.df$Batch,
                          levels=c("SMART", "droplet"),
                          labels=c("SMARTseq", "Droplet"))

ggplot(umap.df, aes(x=UMAP1, y=UMAP2)) +
  geom_point_rast(data=umap.df[, c("UMAP1", "UMAP2")],
                  colour='grey80', size=0.1) +
  geom_point_rast(aes(colour=Dataset), size=0.1) +
  theme_cowplot() +
  scale_colour_aaas() +
  facet_wrap(~Dataset) +
  guides(colour=guide_legend(override.aes=list(size=3))) +
  theme(strip.background=element_rect(colour='white', fill='white'),
        strip.text=element_blank()) +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_MNN-UMAP.pdf",
         height=2.55, width=6.15, useDingbats=FALSE) +
  NULL
```


```{r, warning=FALSE, message=FALSE}
thymus.groups <- read.table("~/Dropbox/Milo/simulations/Thymus_NhoodGroups.txt",
                            sep="\t", header=TRUE, stringsAsFactors=FALSE)
table(thymus.groups$Nhood.Group)
```

For the MNN label transfer, we only need to visualise the nhood groups for the Intertypical TEC MNNs (groups 3,4 & 5).

```{r, warning=FALSE, message=FALSE, fig.height=4.15, fig.width=1.95}
nhood.groups <- unique(thymus.groups$Nhood.Group)
group.cols <- pal_d3()(9)
names(group.cols) <- paste0("Nhood Group ", c(1:9))

umap.df$Nhood.Group <- NA
for(i in seq_along(nhood.groups)){
  i.cells <- thymus.groups$CellID[thymus.groups$Nhood.Group %in% nhood.groups[i]]
  umap.df$Nhood.Group[umap.df$CellID %in% i.cells] <- nhood.groups[i]
}

umap.df$Nhood.Group <- factor(umap.df$Nhood.Group,
                              levels=c(1:9),
                              labels=paste0("Nhood Group ", c(1:9)))

ggplot(umap.df[umap.df$Nhood.Group %in% paste0("Nhood Group ", c(3:5)), ],
       aes(x=UMAP1, y=UMAP2)) +
  geom_point_rast(data=umap.df[, c("UMAP1", "UMAP2")],
                  colour='grey80', size=0.1) +
  geom_point_rast(data=umap.df[umap.df$Nhood.Group %in% paste0("Nhood Group ", c(3:5)), ],
                  aes(colour=Nhood.Group), size=0.5) +
  theme_cowplot() +
  scale_colour_manual(values=group.cols) +
  facet_wrap(~Nhood.Group, ncol=1) +
  theme(strip.background=element_rect(colour='white', fill='white'),
        strip.text=element_text(size=8, face='bold'),
        axis.title=element_text(size=8),
        axis.text=element_text(size=8)) +
  guides(colour=FALSE) +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_NhoodGroup_MNNUMAP-select.pdf",
         height=4.15, width=1.95, useDingbats=FALSE) +
  NULL
```

We also generate the whole plot for all nhoods for transparency.

```{r, warning=FALSE, message=FALSE, fig.height=5.15, fig.width=5.95}
ggplot(umap.df[!is.na(umap.df$Nhood.Group), ],
       aes(x=UMAP1, y=UMAP2)) +
  geom_point_rast(data=umap.df[, c("UMAP1", "UMAP2")],
                  colour='grey80', size=0.1) +
  geom_point_rast(data=umap.df[!is.na(umap.df$Nhood.Group), ],
                  aes(colour=Nhood.Group), size=0.5) +
  theme_cowplot() +
  scale_colour_manual(values=group.cols) +
  facet_wrap(~Nhood.Group) +
  theme(strip.background=element_rect(colour='white', fill='white'),
        strip.text=element_text(size=8, face='bold'),
        axis.title=element_text(size=8),
        axis.text=element_text(size=8)) +
  guides(colour=FALSE) +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_NhoodGroup_MNNUMAP.pdf",
         height=5.15, width=5.95, useDingbats=FALSE) +
  NULL
```




```{r, warning=FALSE, message=FALSE, fig.height=1.95, fig.width=1.95}
set.seed(42)
random.cell <- sample(umap.df$CellID[umap.df$Nhood.Group %in% c("Nhood Group 3")], size=1)
umap.df$Rand.Cell <- NA
umap.df$Rand.Cell[umap.df$CellID %in% random.cell] <- "Select"

ggplot(umap.df[!is.na(umap.df$Nhood.Group), ], aes(x=UMAP1, y=UMAP2)) +
  geom_point_rast(data=umap.df[, c("UMAP1", "UMAP2")],
                  colour='grey80', size=0.1) +
  geom_point_rast(data=umap.df[umap.df$Nhood.Group %in% c("Nhood Group 3"), ],
                  aes(colour=Rand.Cell), size=0.5) +
  theme_cowplot() +
  scale_colour_manual(values=c("#3B4992FF")) +
  theme(strip.background=element_rect(colour='white', fill='white'),
        strip.text=element_text(size=8, face='bold'),
        axis.title=element_text(size=8),
        axis.text=element_text(size=8)) +
  guides(colour=FALSE) +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_MNNUMAP-randCell.pdf",
         height=1.95, width=1.95, useDingbats=FALSE) +
  NULL
```

Now we find the MNNs for this cell, and plot those on the UMAP.

```{r, warning=FALSE, message=FALSE}
mnn.pair.list <- readRDS("~/Dropbox/Milo/simulations/MNNList.RDS")
```

```{r, warning=FALSE, message=FALSE, fig.height=1.95, fig.width=1.95}
set.seed(42)
rand.cell.nhood <- gsub(as.character(umap.df$Nhood.Group[umap.df$Rand.Cell %in% c("Select")]),
                        pattern="Nhood Group ", replacement="")
rand.mnn <- mnn.pair.list[[rand.cell.nhood]]
rand.mnn.cells <- rand.mnn$CellID.Second[rand.mnn$CellID.First %in% random.cell]

umap.df$Rand.MNN <- NA
umap.df$Rand.MNN[umap.df$CellID %in% rand.mnn.cells] <- "MNN"

ggplot(umap.df[!is.na(umap.df$Nhood.Group), ], aes(x=UMAP1, y=UMAP2)) +
  geom_point_rast(data=umap.df[, c("UMAP1", "UMAP2")],
                  colour='grey80', size=0.1) +
  geom_point_rast(data=umap.df,
                  aes(colour=Rand.MNN), size=0.5) +
  theme_cowplot() +
  scale_colour_manual(values=c("#EE0000FF")) +
  theme(strip.background=element_rect(colour='white', fill='white'),
        strip.text=element_text(size=8, face='bold'),
        axis.title=element_text(size=8),
        axis.text=element_text(size=8)) +
  guides(colour=FALSE) +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_MNNUMAP-randMNNs.pdf",
         height=1.95, width=1.95, useDingbats=FALSE) +
  NULL
```

```{r, warning=FALSE, message=FALSE, fig.height=1.95, fig.width=1.95}
mnn.list <- readRDS("~/Dropbox/Milo/simulations/nhood_mnn_list.RDS")

umap.df$MNN.NhoodGroup <- NA
for(i in seq_along(nhood.groups)){
  i.cells <- mnn.list[[paste0(nhood.groups[i])]]
  umap.df$MNN.NhoodGroup[umap.df$CellID %in% i.cells] <- nhood.groups[i]
}

umap.df$MNN.NhoodGroup <- factor(umap.df$MNN.NhoodGroup,
                                 levels=c(1:9),
                                 labels=paste0("Nhood Group ", c(1:9)))

ggplot(umap.df[!is.na(umap.df$MNN.NhoodGroup), ],
       aes(x=UMAP1, y=UMAP2)) +
  geom_point_rast(data=umap.df[, c("UMAP1", "UMAP2")],
                  colour='grey80', size=0.1) +
  geom_point_rast(data=umap.df[!is.na(umap.df$Rand.MNN) & umap.df$MNN.NhoodGroup %in% c("Nhood Group 3"), ],
                  aes(colour=MNN.NhoodGroup), size=0.5) +
  theme_cowplot() +
  scale_colour_manual(values=group.cols) +
  theme(strip.background=element_rect(colour='white', fill='white'),
        strip.text=element_text(size=8, face='bold'),
        axis.title=element_text(size=8),
        axis.text=element_text(size=8)) +
  guides(colour=FALSE) +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_MNNUMAP-randMNN-NhoodGroup.pdf",
         height=1.95, width=1.95, useDingbats=FALSE) +
  NULL
```




Now we do a label transfer of the neighbourhood groups from the SMART-seq data onto the droplet data.

```{r, warning=FALSE, message=FALSE, fig.height=3.95, fig.width=4.15}


ggplot(umap.df[!is.na(umap.df$MNN.NhoodGroup), ], aes(x=UMAP1, y=UMAP2)) +
  geom_point_rast(data=umap.df[, c("UMAP1", "UMAP2")],
                  colour='grey80', size=0.1) +
  geom_point_rast(data=umap.df[!is.na(umap.df$MNN.NhoodGroup), ],
                  aes(colour=MNN.NhoodGroup), size=0.5) +
  theme_cowplot() +
  scale_colour_manual(values=group.cols) +
  facet_wrap(~MNN.NhoodGroup) +
  theme(strip.background=element_rect(colour='white', fill='white'),
        strip.text=element_text(size=8, face='bold'),
        axis.title=element_text(size=8),
        axis.text=element_text(size=8)) +
  guides(colour=FALSE) +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_NhoodGroupDroplet_MNNUMAP.pdf",
         height=3.95, width=4.15, useDingbats=FALSE) +
  NULL
```

I have pre-computed the proportions of these MNNs over samples and ages.

```{r, warning=FALSE, message=FALSE}
mnn.props <- read.table("~/Dropbox/Milo/simulations/Thymus-MNN_props.txt",
                        sep="\t", header=TRUE, stringsAsFactors=FALSE)
colnames(mnn.props) <- c("SampID", "Age", "Freq", "NhoodGroup")
mnn.props$Age <- factor(mnn.props$Age,
                        levels=c("Wk1", "Wk4", "Wk16"),
                        labels=c("1wk", "4wk", "16wk"))

mnn.props$NhoodGroup <- ordered(mnn.props$NhoodGroup,
                                levels=c(1:9),
                                labels=paste0("Nhood Group ", c(1:9)))
```

From our previous analyses, we have seen that the mTEC-biased Intertypical TEC correspond to Nhood group 5. The others correspond to the enriched 
Intertypical TEC and the mature mTEC and proliferating TEC.

```{r, warning=FALSE, message=FALSE, fig.width=5.25, fig.height=4.25}
set.seed(42)
age.cols <- viridis(option="magma", n=5)
names(age.cols) <- c("4wk", "32wk", "1wk", "16wk", "52wk")

ggplot(mnn.props[mnn.props$NhoodGroup %in% paste0("Nhood Group ", c(1:9)), ],
       aes(x=Age, y=Freq)) +
  geom_boxplot(aes(fill=Age), alpha=0.5) +
  geom_jitter(aes(colour=Age), position=position_jitter(width=0.1)) +
  theme_cowplot() +
  facet_wrap(~NhoodGroup, scales="free", ncol=3) +
  expand_limits(y=c(0)) +
  # coord_flip() +
  scale_fill_manual(values=age.cols, labels=c("1wk", "4wk", "16wk", "32wk", "52wk")) +
  scale_colour_manual(values=age.cols, labels=c("1wk", "4wk", "16wk", "32wk", "52wk")) +
  theme(#aspect=1,
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y=element_text(size=9),
        axis.title.y=element_text(size=9),
        strip.text=element_text(size=9, face='bold'),
        strip.background=element_rect(fill='white', colour='white')) +
  labs(y="Proportion", x="") +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_dropletMNN-boxplots.pdf",
         height=4.25, width=5.25, useDingbats=FALSE) +
  NULL
```



```{r, warning=FALSE, message=FALSE, fig.width=1.75, fig.height=4.15}
set.seed(42)
ggplot(mnn.props[mnn.props$NhoodGroup %in% paste0("Nhood Group ", c(3:5)), ],
       aes(x=Age, y=Freq)) +
  geom_boxplot(aes(fill=Age), alpha=0.5) +
  geom_jitter(aes(colour=Age), position=position_jitter(width=0.1)) +
  theme_cowplot() +
  facet_wrap(~NhoodGroup, scales="free", ncol=1) +
  expand_limits(y=c(0)) +
  # coord_flip() +
  scale_fill_manual(values=age.cols, breaks=c("Wk1", "Wk4", "Wk16")) +
  scale_colour_manual(values=age.cols, breaks=c("Wk1", "Wk4", "Wk16")) +
  theme(#aspect=1,
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y=element_text(size=9),
        axis.title.y=element_text(size=10, face='bold'),
        strip.text=element_text(size=9, face='bold'),
        strip.background=element_rect(fill='white', colour='white')) +
  labs(y="Proportion", x="") +
  guides(fill=FALSE,
         colour=FALSE) +
  ggsave("~/Dropbox/Milo/revision/figures/MouseThymus_dropletMNN-boxplots_select.pdf",
         height=4.15, width=1.75, useDingbats=FALSE) +
  NULL
```

