---
title: "Mouse thymus: Validating differential abundance findings"
output: html_notebook
---

# Introduction

In Baran-Gale _et al._, we validated our finding that accumulating progenitors is coincident with a restriction in mTEC differentiation. These results with Milo 
suggest that the mTEC-biased progenitors are where the defect manifests. To validate this we can look at the similarity to cells from our larger droplet 
scRNAseq experiment, and how the abundance of these cells changes across ages.

I will pull in the 10X data from our public-facing FTP server, which I've then organised into a `SingleCellExperiment` object for convenience, and compute the 
distance from the cells in the DA nhoods to the cells in the larger data set. Then compute the abundance of these cells in each replicated sample and plot how 
they change across ages.

```{r, message=FALSE}
library(batchelor)
library(SingleCellExperiment)
library(miloR)
library(cowplot)
library(ggplot2)
library(igraph)
library(ggthemes)
library(biomaRt)
library(ggsci)
library(irlba)
library(umap)
library(viridis)
library(ComplexHeatmap)
library(ggbeeswarm)
library(colorspace)
library(patchwork)
library(enrichR)
library(scater)
library(reshape2)
library(BiocNeighbors)
```

```{r, defining-nhood-groups, warning=FALSE, message=FALSE}
thymus.groups <- read.table("/nfs/research1/marioni/mdmorgan/milo_testing/results/Thymus_NhoodGroups.txt",
	                    sep="\t", header=TRUE, stringsAsFactors=FALSE)
rownames(thymus.groups) <- thymus.groups$CellID
head(thymus.groups)
```

I have already batch-corrected the SMART-seq and 10X droplet data together. All that remains is to find the set of 
mutual nearest neighbours to the DA nhood cells, and compute their proportions across ages.


```{r, load-big-SCE, warning=FALSE, message=FALSE}
big.sce <- readRDS("/nfs/research1/marioni/mdmorgan/milo_testing/data/CombinedThymus_SCE.RDS")
```

### Computing distances

First I will find the K-mutual nearest neighbours for each cell in the different nhood groups. Then as a group, I will compute their abundance within each experimental sample.

```{r, Finding-MNNs, warning=FALSE, message=FALSE}
nhood.group.mnn <- findMutualNN(data1=reducedDim(big.sce[,big.sce$Batch %in% c("SMART")], "MNN"),
		                data2=reducedDim(big.sce[,big.sce$Batch %in% c("droplet")], "MNN"),
				k1=20, k2=60)
```

After computing the MNNs between the 2 data sets, we will subset the MNNs of the DA nhood cells.

```{r, Extracting-NhoodGroup-MNNs, warning=FALSE, message=FALSE}
nhood.groups <- unique(thymus.groups$Nhood.Group)

nhood.mnn.list <- list()
for(x in seq_along(nhood.groups)){
  x.group <- nhood.groups[x]
  x.cells <- rownames(thymus.groups[thymus.groups$Nhood.Group %in% x.group, ])
  x.idx <- which(colnames(big.sce[, big.sce$Batch %in% c("SMART")]) %in% x.cells)
  if(length(x.idx) > 0){
    x.mnn.idx <- which(nhood.group.mnn$first %in% x.idx)
    x.mnn.first <- nhood.group.mnn$first[x.mnn.idx]
    x.mnn.second <- nhood.group.mnn$second[x.mnn.idx]
    x.mnn.cells <- colnames(big.sce[, big.sce$Batch %in% c("droplet")])[x.mnn.second]
    nhood.mnn.list[[paste0(x.group)]] <- x.mnn.cells
  } else{
    nhood.mnn.list[[paste0(x.group)]] <- c()
  }
}
```

Having now extracted all of the MNNs, we can tally these up to the relevant meta-data and compute their proportions w.r.t. age.

```{r, load-droplets, warning=FALSE, message=FALSE}
droplet.sce <- readRDS("/nfs/research1/marioni/mdmorgan/milo_testing/data/ThymusDroplet_SCE.RDS")
```

```{r, Computing-MNN-proportions, warning=TRUE, message=TRUE}
samp.nums <- table(colData(droplet.sce)$SampID)
mnn.props.list <- list()
for(x in seq_along(nhood.groups)){
  x.group <- nhood.groups[x]
  print(x.group)
  x.cells <- nhood.mnn.list[[paste0(x.group)]]      
  print(length(x.cells))
  x.nums <- table(colData(droplet.sce)[x.cells, ]$SampID, colData(droplet.sce)[x.cells, ]$Age)
  
  if(nrow(x.nums) > 0){
    print(x.nums)
    print(dim(x.nums))
    #x.props <- as.data.frame(apply(x.nums, 1, FUN=function(X) X/samp.nums))
    #print(head(x.props))
    #x.props$SampID <- colnames(x.nums)
    #x.melt <- melt(x.props, id.vars='SampID')
    #print(head(x.props))
    #if(nrow(x.props) > 0){
    #  x.props$Nhood.Group <- x.group
    # mnn.props.list[[paste0(x.group)]] <- x.props
    #}
  }
}

mnn.prop.df <- do.call(rbind.data.frame, mnn.props.list)
write.table(mnn.prop.df, file="/nfs/research1/marioni/mdmorgan/milo_testing/results/Thymus-MNN_props.txt", sep="\t", quote=FALSE, row.names=FALSE)

head(mnn.prop.df)
```

