---
title: "Batch effect simulation"
output: html_notebook
---

```{r}
library(miloR)
library(dyntoy)
library(tidyverse)
```

We want to check the effect of batch effects on Milo DA testing in a controlled setting, using simulations

## Make simulated dataset
```{r, warning=FALSE, message=FALSE}
set.seed(42)
dataset <- generate_dataset(
  model = model_linear(num_milestones = 10),
  num_cells = 5000,
  num_features = 5000
)

gex <- as.matrix(dataset$counts) ## <---- HERE CHANGE TO COUNTS
branches <- dataset$prior_information$groups_id
  
```

I assign cells to simulated biological conditions and replicates, that we will use for differential abundance testing. For each of the $M$ clusters, I assign different proportions of cells to condition A or condition B, while simulating proportionate mixing between replicates.

```{r}
coldata_df <- data.frame(cell_id = rownames(gex))
coldata_df <- left_join(coldata_df, branches)
  
## Simulate DA between conditions
prob_start <- 0.05
prob_end <- 0.95
n_groups <- length(unique(branches$group_id))
da_groups <- c("M8", "M9", "M10")
p_vec <- ifelse(paste0("M", 1:10) %in% da_groups, 0.2, 0.5)
# p_vec <- seq(prob_start, prob_end, length.out = n_groups)
a.cells <- c()
for (i in 1:n_groups) {
  g <- paste0("M",i)
  p <- p_vec[i] 
  m.A <- sample(coldata_df$cell_id[coldata_df$group_id==g], 
                size=floor(sum(coldata_df$group_id==g)*p))
  a.cells <- c(a.cells, m.A)
}

coldata_df <- coldata_df %>% dplyr::mutate(condition = ifelse(cell_id %in% a.cells, "A", 'B')) 

## Simulate replicates
coldata_df <- coldata_df %>%
  group_by(group_id) %>%
  dplyr::mutate(replicate=c(rep("R1", floor(n()*0.3)), 
                            rep("R2", floor(n()*0.3)), 
                            rep("R3", n() - 2*(floor(n()*0.3))))
  ) 

## Add sample name (condition + replicate)
coldata_df$sample <- paste(coldata_df$condition, coldata_df$replicate, sep="_")

coldata_df %>%
  group_by(group_id, condition) %>%
  summarise(n=n()) %>%
  pivot_wider(id_cols=group_id, names_from=condition, values_from=n)

```

Finally I can construct a `Milo` object for DA analysis.

```{r}
## Make SingleCellExperiment object
coldata <- coldata_df %>% column_to_rownames("cell_id")
sim_sce <- SingleCellExperiment(assay=list(counts=t(gex)), colData=coldata)
logcounts(sim_sce) <- log2(counts(sim_sce) + 1)
sim_sce <- scater::runPCA(sim_sce, ncomponents=30, scale=TRUE)

## Make milo object
sim_milo <- Milo(sim_sce)
sim_milo
```


## Add random batch effect to the replicates

```{r}
sim_milo_be <- sim_milo
logcounts(sim_milo_be) <- log2(counts(sim_milo_be) + 1)

assay(sim_milo_be, 'scaled_logcounts') <- scale(logcounts(sim_milo_be))

set.seed(2020)
batch_effect1 <- rnorm(n = nrow(sim_milo), sd=0.5, mean=0)
set.seed(2022)
batch_effect2 <- rnorm(n = nrow(sim_milo), sd=0.5, mean=0)

for (smp in unique(sim_milo_be$sample)){
  if (smp %in% c("A_R1", "B_R1")) {
    assay(sim_milo_be, 'scaled_logcounts')[,sim_milo_be$sample == smp] <- 
      apply(assay(sim_milo_be, 'scaled_logcounts')[,sim_milo_be$sample == smp], 2, function(x) x + batch_effect1)
  } else {
    assay(sim_milo_be, 'scaled_logcounts')[,sim_milo_be$sample == smp] <- 
      apply(assay(sim_milo_be, 'scaled_logcounts')[,sim_milo_be$sample == smp], 2, function(x) x + batch_effect2)
    }
  }

logcounts(sim_milo_be) <- assay(sim_milo_be, 'scaled_logcounts') 
sim_milo_be <- scater::runPCA(sim_milo_be, ncomponents=30, scale=FALSE)
sim_milo_be <- runUMAP(sim_milo_be, pca=30)

## Save as batch covariate
colData(sim_milo_be)['technical_batch'] <- ifelse(sim_milo_be$sample %in% c("A_R1", "B_R1"), "B1", "B2")
```

```{r}
plotUMAP(sim_milo_be, colour_by="sample") + plotUMAP(sim_milo_be, colour_by="technical_batch")
```

## Run MNN correction

```{r}
library(batchelor)
# sim_milo_be_mnn <- sim_milo_be

## Split in two SCE objects
b1_milo <- sim_milo_be[,sim_milo_be$technical_batch=="B1"]
b2_milo <- sim_milo_be[,sim_milo_be$technical_batch=="B2"]

sim_milo_be_mnn <- batchelor::fastMNN(b1_milo, b2_milo)
sim_milo_be_mnn <- runUMAP(sim_milo_be_mnn, dimred="corrected")
```

```{r, fig.width=10, fig.height=4}
## Reorder cells and add coldata
sim_milo_be_mnn <- sim_milo_be_mnn[,colnames(sim_milo_be)]
colData(sim_milo_be_mnn) <- colData(sim_milo_be)

plotUMAP(sim_milo_be_mnn, colour_by="sample") + plotUMAP(sim_milo_be_mnn, colour_by="technical_batch")
```
```{r}
plotUMAP(sim_milo_be_mnn, colour_by="group_id", text_by="group_id")
plotUMAP(sim_milo_be_mnn, colour_by="condition")
```
Make good plots
```{r, fig.height=8, fig.width=7}
umap_mnn_df <- data.frame(reducedDim(sim_milo_be_mnn,"UMAP"))
colnames(umap_mnn_df) <- c("UMAP1", "UMAP2") 
umap_mnn_df[,"UMAP1"] <- jitter(umap_mnn_df[,"UMAP1"], factor=100)
umap_mnn_df[,"UMAP2"] <- jitter(umap_mnn_df[,"UMAP2"], factor=100)
umap_mnn_df <- cbind(umap_mnn_df, data.frame(colData(sim_milo_be_mnn)[c("technical_batch", "condition", "group_id")])) %>%
  mutate(trueDA=ifelse(group_id %in% da_groups, "is DA", NA))  %>%
  mutate(technical_batch=factor(technical_batch, levels=c("B2", "B1")))

umap_uncor_df <- data.frame(reducedDim(sim_milo_be,"UMAP"))
colnames(umap_uncor_df) <- c("UMAP1", "UMAP2") 
umap_uncor_df[,"UMAP1"] <- jitter(umap_uncor_df[,"UMAP1"], factor=100)
umap_uncor_df[,"UMAP2"] <- jitter(umap_uncor_df[,"UMAP2"], factor=100)
umap_uncor_df <- cbind(umap_uncor_df, data.frame(colData(sim_milo_be)[c("technical_batch", "condition", "group_id")])) %>%
  mutate(trueDA=ifelse(group_id %in% da_groups, "is DA", NA)) %>%
  mutate(technical_batch=factor(technical_batch, levels=c("B2", "B1")))

mnn_1 <- ggplot(umap_mnn_df, aes(UMAP1, UMAP2)) +
  geom_point(size=0.1, alpha=0.1) +
  geom_point(data=.%>% dplyr::filter(!is.na(trueDA)), aes(color="is DA"), size=0.3) +
  # coord_fixed() +
  scale_color_manual(values="red", name="") +
  theme_classic(base_size = 20) +
  guides(color=guide_legend(override.aes = list(size=2))) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.position = "top", plot.title=element_text(hjust=0.5)) +
  # ggtitle("MNN corrected")
  NULL

mnn_2 <- ggplot(umap_mnn_df %>% arrange(UMAP1), aes(UMAP1, UMAP2, color=technical_batch)) +
  geom_point(size=0.3, alpha=1) +
  # coord_fixed() +
  scale_color_brewer(palette="Set2", name="Batch") +
  theme_classic(base_size = 20) +
  guides(color=guide_legend(override.aes = list(size=2))) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.position = "top", plot.title=element_text(hjust=0.5)) 
  # ggtitle("MNN corrected")

uncor_1 <- ggplot(umap_uncor_df, aes(UMAP1, UMAP2)) +
  geom_point(size=0.1, alpha=0.1) +
  geom_point(data=.%>% dplyr::filter(!is.na(trueDA)), aes(color="is DA"), size=0.3) +
  # coord_fixed() +
  scale_color_manual(values="red", name="") +
  theme_classic(base_size = 20) +
  guides(color=guide_legend(override.aes = list(size=2))) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.position = "top", plot.title=element_text(hjust=0.5)) +
  # ggtitle("MNN corrected")
  NULL

uncor_2 <- ggplot(umap_uncor_df %>% arrange(sample(nrow(umap_uncor_df))), aes(UMAP1, UMAP2, color=technical_batch)) +
  geom_point(size=0.3, alpha=1) +
  # coord_fixed() +
  scale_color_brewer(palette="Set2", name="Batch") +
  theme_classic(base_size = 20) +
  # facet_grid("Uncorrected"~., switch = 'y') +
  guides(color=guide_legend(override.aes = list(size=2))) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        # strip.placement = "outside",
        legend.background = element_blank(),
        legend.position = "top", plot.title=element_text(hjust=0.5)) 
  # ggtitle("MNN corrected")


pan_1 <- ((uncor_2  + uncor_1) / ((mnn_2 + guides(color="none")) + (mnn_1 + guides(color="none")))) 
pan_1
```

## Run DA analysis with different designs

Make design matrix

```{r}
sim_design <- data.frame(colData(sim_milo_be))[,c("sample", "condition", "technical_batch")]
sim_design <- distinct(sim_design)
rownames(sim_design) <- NULL
sim_design <- column_to_rownames(sim_design, "sample")
sim_design
```


```{r}

## Milo pre-processing
k=10

sim_milo_be <- buildGraph(sim_milo_be, k = k, d = 30)
sim_milo_be <- makeNhoods(sim_milo_be, k = k, d = 30, prop = 0.1, refined = TRUE)
sim_milo_be <- countCells(sim_milo_be , meta.data = data.frame(colData(sim_milo_be)), sample="sample")
sim_milo_be <- calcNhoodDistance(sim_milo_be, d = 30)

reducedDim(sim_milo_be_mnn, "PCA") <- reducedDim(sim_milo_be_mnn, "corrected")
sim_milo_be_mnn <- buildGraph(sim_milo_be_mnn, k = k, d = 30)
sim_milo_be_mnn <- makeNhoods(sim_milo_be_mnn, k = k, d = 30, prop = 0.1, refined = TRUE)
sim_milo_be_mnn <- countCells(sim_milo_be_mnn , meta.data = data.frame(colData(sim_milo_be)), sample="sample")
sim_milo_be_mnn <- calcNhoodDistance(sim_milo_be_mnn, d = 30)

## Run test
da_res_be_cond <- testNhoods(sim_milo_be, design = ~ condition, design.df = sim_design)
da_res_be_cond_tech <- testNhoods(sim_milo_be, design = ~ technical_batch + condition, design.df = sim_design)
da_res_mnn_cond <- testNhoods(sim_milo_be_mnn, design = ~ condition, design.df = sim_design)
da_res_mnn_cond_tech <- testNhoods(sim_milo_be_mnn, design = ~ technical_batch + condition, design.df = sim_design)

```

Diagnostic plots

```{r, fig.height=5, fig.width=10}
ggplot(da_res_be_cond, aes(PValue)) + geom_histogram() +
ggplot(da_res_be_cond, aes(logFC, -log10(SpatialFDR))) + geom_point() +
  ggtitle("uncorrected - design = ~ condition")

ggplot(da_res_be_cond_tech, aes(PValue)) + geom_histogram() +
ggplot(da_res_be_cond_tech, aes(logFC, -log10(SpatialFDR))) +
  geom_point() +
  ggtitle("uncorrected - design = ~ batch + condition")

ggplot(da_res_mnn_cond, aes(PValue)) + geom_histogram() +
ggplot(da_res_mnn_cond, aes(logFC, -log10(SpatialFDR))) +
  geom_point() +
  ggtitle("MNN corrected - design = ~ condition")

ggplot(da_res_mnn_cond_tech, aes(PValue)) + geom_histogram() +
ggplot(da_res_mnn_cond_tech, aes(logFC, -log10(SpatialFDR))) +
  geom_point() +
  ggtitle("MNN corrected - design = ~ batch + condition")


```


```{r}
myPlotNhoodGraph <- function(x, layout="UMAP", colour_by=NA, subset.nhoods=NULL, jitter=TRUE, ... ){
  ## Check for valid nhoodGraph object
  if (is.character(layout)) {
    if (!layout %in% names(reducedDims(x))) {
      stop(paste(layout, "isn't in readucedDim(x) - choose a different layout"))
    }
  }
  nh_graph <- nhoodGraph(x)

  ## Subset
  if (!is.null(subset.nhoods)) {
    nh_graph <- igraph::induced_subgraph(nh_graph, vids = which(as.numeric(V(nh_graph)$name) %in% unlist(nhoodIndex(x)[subset.nhoods])))
  }


  # ## Order vertex ids by size (so big nhoods are plotted first)
  # nh_graph <- permute(nh_graph, order(vertex_attr(nh_graph)$size))

  ## Define node color
  if (!is.na(colour_by)) {
    if (colour_by %in% colnames(colData(x))) {

      col_vals <- colData(x)[as.numeric(vertex_attr(nh_graph)$name), colour_by]
      if (!is.numeric(col_vals)) {
        col_vals <- as.character(col_vals)
        }
      V(nh_graph)$colour_by <- col_vals
    } else {
      stop(paste(colour_by, "is not a column in colData(x)"))
    }
  } else {
    V(nh_graph)$colour_by <- V(nh_graph)$size
    colour_by <- "Nhood size"
  }
  # print(order(abs(vertex_attr(nh_graph)$colour_by)))
  # before <- vertex_attr(nh_graph, "name")
  
  # after <- vertex_attr(nh_graph, "name")
  # print(before==after)
  
  if (jitter){
    set.seed(42)
    reducedDim(x, layout)[as.numeric(vertex_attr(nh_graph)$name),1] <- jitter(reducedDim(x, layout)[as.numeric(vertex_attr(nh_graph)$name),1], factor=400)
    reducedDim(x, layout)[as.numeric(vertex_attr(nh_graph)$name),2] <- jitter(reducedDim(x, layout)[as.numeric(vertex_attr(nh_graph)$name),2], factor=400)
    # layout[,1] <- jitter(layout[,1], factor=400)
    # layout[,2] <- jitter(layout[,2], factor=400)
  }
  
  nh_graph <- igraph::permute(nh_graph, order(abs(vertex_attr(nh_graph)$colour_by), decreasing = FALSE))
  
  ## Define layout
  if (is.character(layout)) {
    redDim <- layout
    layout <- reducedDim(x, redDim)[as.numeric(vertex_attr(nh_graph)$name),]
    # make sure this is a matrix!
    if(!any(class(layout) %in% c("matrix"))){
        warning("Coercing layout to matrix format")
        layout <- as(layout, "matrix")
    }
  }

  
  
  pl <- ggraph(igraph::simplify(nh_graph), layout = layout) +
    geom_edge_link0(aes(width = weight), edge_colour = "grey66", edge_alpha=0.2) +
    geom_node_point(aes(fill = colour_by, size = size), color="grey20", shape=21) +
    scale_size(range = c(0.5,2), name="Nhood size") +
    scale_edge_width(range = c(0.2,3), name="overlap size") +
    theme_classic(base_size=20) +
    theme(axis.line = element_blank(), axis.text = element_blank(),
          axis.ticks = element_blank(), axis.title = element_blank())
    # theme_graph()

  if (is.numeric(V(nh_graph)$colour_by)) {
    pl <- pl + scale_fill_gradient2(name=colour_by)
  } else {
    pl <- pl + scale_fill_brewer(palette="Spectral", name=colour_by)
  }

  pl
  }


myPlotNhoodGraphDA <- function(x, milo_res, alpha=0.05, ... ){
  if (is.character(layout)) {
    if (!layout %in% names(reducedDims(x))) {
      stop(paste(layout, "is not in readucedDim(x) - choose a different layout"))
    }
  }

  ## Add milo results to colData
  signif_res <- milo_res
  signif_res[signif_res$SpatialFDR > alpha,"logFC"] <- 0
  colData(x)["logFC"] <- 0
  colData(x)[unlist(nhoodIndex(x)[signif_res$Nhood]),] <- signif_res$logFC

  ## Plot logFC
  myPlotNhoodGraph(x, colour_by = "logFC", ... )
}

```


```{r, fig.height=9, fig.width=9}
sim_milo_be <- buildNhoodGraph(sim_milo_be)
sim_milo_be_mnn <- buildNhoodGraph(sim_milo_be_mnn)

nh1 <- myPlotNhoodGraphDA(sim_milo_be, da_res_be_cond, alpha = 0.1) +
  scale_fill_gradient2(limits=c(-2, 5), name="log-Fold Change") +
  scale_size(limits=c(10,130), range = c(0.5,3), name="Nhood size") +
  # coord_fixed() +
  scale_edge_width(range = c(0.2,3), limits=c(0,40), name="overlap size") 
nh2 <- myPlotNhoodGraphDA(sim_milo_be, da_res_be_cond_tech, alpha = 0.1) +
  scale_fill_gradient2(limits=c(-2, 5), name="log-Fold Change") +
  # coord_fixed() +
  scale_size(limits=c(10,130), range = c(0.5,3), name="Nhood size") +
  scale_edge_width(range = c(0.2,3), limits=c(0,40), name="overlap size") 
nh3 <- myPlotNhoodGraphDA(sim_milo_be_mnn, da_res_mnn_cond, alpha = 0.1) +
  scale_fill_gradient2(limits=c(-2, 5), name="log-Fold Change") +
  # coord_fixed() +
  scale_size(limits=c(10,130), range = c(0.5,3), name="Nhood size") +
  scale_edge_width(range = c(0.2,3), limits=c(0,40), name="overlap size") 
nh4 <- myPlotNhoodGraphDA(sim_milo_be_mnn, da_res_mnn_cond_tech, alpha = 0.1) +
  scale_fill_gradient2(limits=c(-2, 5), name="log-Fold Change") +
  # coord_fixed() +
  scale_size(limits=c(10,130), range = c(0.5,3), name="Nhood size") +
  scale_edge_width(range = c(0.2,3), limits=c(0,40), name="overlap size") 

pan_2 <- ((nh1 + facet_wrap("~ condition"~.) ) | nh2 + facet_wrap('~ batch + condition'~.))  /
  ((nh3 ) | nh4) +
  plot_layout(guides="collect") &
  guides(fill=guide_colourbar(title.position="top", title.hjust = 0.5, barwidth = 6),
         size=guide_legend(title.position="top", title.hjust = 0.5, ncol=2),
         edge_width=guide_legend(title.position="top", title.hjust = 0.5, ncol=2)) &
  theme(legend.position = "top")

pan_2
```

```{r, fig.width=15, fig.height=6}
(pan_1 | pan_2) + plot_layout(widths=c(1,1.2)) +
  ggsave("~/milo_output/batch_effect_sim_umaps.pdf", width = 15, height = 7)
```


<!-- ```{r} -->
<!-- add_nhood_coldata_to_res(sim_milo_be_mnn, da_res_mnn_cond, "group_id") %>% -->
<!--   mutate(color=ifelse(group_id %in% da_groups, TRUE, FALSE)) %>% -->
<!--   ggplot( aes(logFC, -log10(SpatialFDR), color=color)) + -->
<!--   geom_point() + -->
<!--   ggtitle("MNN corrected - design = ~ condition") -->

<!-- ``` -->



## Compare FPR / TPR

```{r}
## Add if cells are TP or TN
true_DA <- ifelse(sim_milo_be$group_id %in% da_groups, "TP", "TN")
colData(sim_milo_be)["trueDA"] <- factor(true_DA) ## Make factor otherwise counting with table doesn't work
true_DA <- ifelse(sim_milo_be_mnn$group_id %in% da_groups, "TP", "TN")
colData(sim_milo_be_mnn)["trueDA"] <- factor(true_DA) ## Make factor otherwise counting with table doesn't work

## Find which nhoods are TP / TN
add_nhood_coldata_to_res <- function(liver_milo, milo_res, coldata_col){
  nhood_counts <- sapply(seq_along(nhoods(liver_milo)), function(x) table(colData(liver_milo)[as.vector(nhoods(liver_milo)[[x]]), coldata_col]))
  nhood_counts <- t(nhood_counts)
  rownames(nhood_counts) <- seq_along(nhoods(liver_milo))
  max_val <- apply(nhood_counts, 1, function(x) colnames(nhood_counts)[which.max(x)])
  max_frac <- apply(nhood_counts, 1, function(x) max(x)/sum(x))
  milo_res[coldata_col] <- max_val
  milo_res[paste0(coldata_col, "_fraction")] <- max_frac
  return(milo_res)
}

da_res_be_cond <- add_nhood_coldata_to_res(sim_milo_be, da_res_be_cond, coldata_col = "trueDA") 
  # mutate(trueDA = ifelse(trueDA=="TP" & trueDA_fraction > 0.7, "TP", "TN"))
da_res_be_cond_tech <- add_nhood_coldata_to_res(sim_milo_be, da_res_be_cond_tech, coldata_col = "trueDA") 
  # mutate(trueDA = ifelse(trueDA=="TP" & trueDA_fraction > 0.7, "TP", "TN"))
da_res_mnn_cond <- add_nhood_coldata_to_res(sim_milo_be_mnn, da_res_mnn_cond, coldata_col = "trueDA") 
  # mutate(trueDA = ifelse(trueDA=="TP" & trueDA_fraction > 0.7, "TP", "TN"))
da_res_mnn_cond_tech <- add_nhood_coldata_to_res(sim_milo_be_mnn, da_res_mnn_cond_tech, coldata_col = "trueDA") 
  # mutate(trueDA = ifelse(trueDA=="TP" & trueDA_fraction > 0.7, "TP", "TN"))
```

```{r}
colData(sim_milo_be_mnn)["group_id"] <- factor(colData(sim_milo_be_mnn)[["group_id"]])
add_nhood_coldata_to_res(sim_milo_be_mnn, da_res_mnn_cond, coldata_col = "group_id")  
da_res_mnn_cond %>%
  # mutate(da=ifelse(group_id %in% da_groups, "isDA", "notDA")) %>%
  ggplot(aes(logFC, -log10(SpatialFDR))) +
  geom_hline(yintercept = -log10(0.1)) +
  geom_point()
```



```{r}
res_ls <- list(uncor_cond = da_res_be_cond,
     uncor_cond_tech= da_res_be_cond_tech,
     mnn_cond = da_res_mnn_cond,
     mnn_cond_tech = da_res_mnn_cond_tech)

outcomes_df <-
  imap(res_ls, ~ .x %>%
    mutate(predDA = ifelse(SpatialFDR < 0.1, "Positive", "Negative")) %>%
    mutate(outcome = case_when(trueDA == "TP" & predDA == "Positive" ~ "TP",
                             trueDA == "TP" & predDA == "Negative" ~ "FN",
                             trueDA == "TN" & predDA == "Positive" ~ "FP",
                             trueDA == "TN" & predDA == "Negative" ~ "TN"
                             )) %>%
    mutate(outcome=factor(outcome, levels = c("TP", "TN", "FP", "FN"))) %>%
    group_by(outcome) %>%
    summarise(n = n()) %>%
    mutate(set=.y)
  ) %>%
  purrr::reduce(bind_rows) 

outcomes_df %>%
  pivot_wider(id_cols='set', names_from=outcome, values_from="n", values_fill = 0) %>%
  mutate(TPR=TP/(TP+FP), FPR=FP/(FP+FN),
         accuracy = (TP + TN)/(TP + TN + FP + FN),
         recall = TP / (TP+FN)) %>%
  mutate(set=str_replace(set, "cond_tech", "~ batch + condition")) %>%
  mutate(set=str_replace(set, "cond$", "~ condition")) %>%
  separate(col=set, sep="_", into=c('dataset', 'design')) %>%
  ggplot(aes(dataset, FPR, fill=design)) +
  geom_col(position="dodge")
```
```{r}
pl_right <- imap(res_ls, ~ .x %>%
    mutate(predDA = ifelse(SpatialFDR < 0.1, "Positive", "Negative")) %>%
    # mutate(outcome=factor(outcome, levels = c("TP", "TN", "FP", "FN"))) %>%
    group_by(predDA, trueDA) %>%
    summarise(n = n()) %>%
    mutate(set=.y)
  ) %>%
  purrr::reduce(bind_rows) %>%
  pivot_wider(id_cols=c('set', "trueDA"), names_from="predDA", values_from="n", values_fill = 0) %>%
  pivot_longer(cols=c("Negative", 'Positive'), names_to="predDA", values_to="n") %>%
  mutate(set=str_replace(set, "cond_tech", "~ batch + condition")) %>%
  mutate(set=str_replace(set, "cond$", "~ condition")) %>%
  separate(col=set, sep="_", into=c('dataset', 'design')) %>%
  mutate(dataset=factor(ifelse(dataset=="mnn", "MNN corrected", "Uncorrected"), levels=c("Uncorrected", "MNN corrected"))) %>%
  mutate(trueDA=factor(ifelse(trueDA=="TP", "is DA", "not DA"), levels=c("not DA", "is DA"))) %>%
  mutate(predDA=factor(ifelse(predDA=="Positive", "is DA", "not DA"), levels=c("not DA", "is DA"))) %>%
  mutate(design=factor(design, levels=c("~ condition", "~ batch + condition"))) %>%
  mutate(outline=ifelse(predDA==trueDA, 1, 0)) %>%
  ggplot(aes(trueDA, predDA)) +
  geom_tile(fill="grey98") +
  geom_text(aes(label=n), size=6) +
  facet_grid(dataset~design) +
  # coord_fixed() +
  ylab("Predicted") +
  xlab("True") +
  # scale_fill_manual(values=c(TRUE))
  scale_fill_gradient(low="grey98", high="#FEE0D2") +
  theme_classic(base_size = 20) +
  guides(
    # fill = guide_colourbar(title.position="top", title.hjust = 0.5, barwidth = 10)
    fill="none"
    ) +
  theme(legend.position = "top", legend.text = element_blank(), legend.justification = 'center') 

pl_right +
  ggsave("~/milo_output/batch_effect_sim_confs.pdf", height = 5.3, width = 5.4)
```

Assemble figure


```{r, fig.width=15, fig.height=6}
(pan_1 | pan_2 | pl_right) + plot_layout(widths=c(1,1.1,0.8)) +
  ggsave("~/milo_analysis_2020/ms/supplement/suppl_figs/suppl_fig1.pdf", width = 18, height = 7.5)
```
  





